\section{Lecture 15 -- 28th November 2024}\label{sec: lecture 15}
We have the following strengthening of the Weierstrass preparation theorem in the cases where the function $f$ lies in $\CC\{z_{1},\dots,z_{n-1}\}[z_{n}]$. 
\begin{proposition}\label{prop: Weierstrass preparation for univariate over convergent power series}
    Let $U\subseteq\CC^{n}$ be open and $D_{R}(0)=D\subseteq\CC$ a disc and $f\in\CC\{z_{1},\dots,z_{n-1}\}[z_{n}]$. If $f:U\times D\to\CC$ is holomorphic of and of order $k$ with respect to the coordinate $w$ on $D$ then there are functions $e(z,w)\in\CC\{z_{1},\dots,z_{n-1}\}[z_{n}]$ holomorphic nonvanishing on $U\times D$ and $\omega(z,w)$ holomorphic on $U\times D$ of the form $w^{k}+a_{k-1}(z)w^{k-1}+\dots+a_{1}(z)w+a_{0}(z)$ such that $f(z,w)=e(z,w)\cdot\omega(z,w)$.
\end{proposition}
\begin{proof}
    By \Cref{thm: Weierstrass preparation,thm: euclidean division for formal power series} we have $f=e\omega=q\omega+r$ so $r=0,e=q$ with uniqueness and $f$ is of the desired form.  
\end{proof}
We now consider some algebraic properties of the ring of convergent power series. 
\begin{theorem}\label{thm: convergent power series is Noetherian}
    The ring of convergent power series $\CC\{z_{1},\dots,z_{n}\}$ is Noetherian. 
\end{theorem}
\begin{proof}
    We proceed by induction on the number of variables $n$. This holds for $n=0$ as $\CC$ is a field and hence Noetherian. 

    Now suppose $\CC\{z_{1},\dots,z_{n-1}\}$ is Noetherian and thus $\CC\{z_{1},\dots,z_{n-1}\}[z_{n}]$ is Noetherian. Consider $\afrak\subseteq\CC\{z_{1},\dots,z_{n-1}\}[z_{n}]$ a nonzero ideal and $f\in\afrak$. By \Cref{lem: order k function}, we can change coordinates such that $f$ is $z_{n}$-regular. Now for $g\in\CC\{z_{1},\dots,z_{n}\}$, \Cref{thm: euclidean division for formal power series} implies that we can write $g=qf+r$ with $r\in\CC\{z_{1},\dots,z_{n-1}\}[z_{n}]$ of degree less than $k$. This yields a homomorphism of $\CC\{z_{1},\dots,z_{n-1}\}$-modules with kernel $f$ given by $\CC\{z_{1},\dots,z_{n}\}\mapsto\CC\{z_{1},\dots,z_{n-1}\}^{\oplus k}$ by $g\mapsto(a_{0}(z'),\dots,a_{k-1}(z'))$ where $r=a_{k-1}(z')z_{n}^{k-1}+\dots+a_{0}(z')$. Let $\overline{\afrak}$ be the image of $\afrak$ under this map. Now noting that $\CC\{z_{1},\dots,z_{n-1}\}^{\oplus k}$ is Noetherian as a module $\overline{\afrak}$ is generated by finitely many elements $\overline{g_{1}},\dots,\overline{g_{s}}$, the images of $g_{1},\dots,g_{s}\in\afrak$. But this shows that $g$ is a linear combination of $g_{1},\dots,g_{s},f$ for $g\in\afrak$ so this forms a finite generating set of the ideal showing that the ring is Noetherian. 
\end{proof}
We can also show that the ring of convergent power series $\CC\{z_{1},\dots,z_{n}\}$ is factorial, that is, a unique factorization domain. Doing so will require the following proposition. 
\begin{proposition}\label{prop: relatively prime germs}
    Let $U\subseteq\CC^{n}$ open containing the origin and $f,g\in\Ocal(U)$. If $f_{0},g_{0}$ are relatively prime in $\Ocal_{0}$ then there is a neighborhood $V$ of the origin such that $f_{z},g_{z}$ are relatively prime for all $z\in V$. 
\end{proposition}
In turn:
\begin{theorem}\label{thm: convergent power series is UFD}
    The ring of convergent power series $\CC\{z_{1},\dots,z_{n}\}$ is a unique factorization domain. 
\end{theorem}
We now discuss meromorphic functions. 
\begin{definition}[Sheaf of Meromorphic Functions]\label{def: sheaf of meromorphic functions}
    The sheaf of meromorphic functions $\Kcal$ on $\CC^{n}$ is given by $\coprod_{z\in\CC^{n}}\Kcal_{z}$ where $\Kcal_{z}$ is the ring of germs of meromorphic functions at $z$. 
\end{definition}
Sections of $\Kcal$ give rise to meromorphic functions in the same way as \Cref{def: continuous section}. 