\section{Prismatic Cohomology}\label{sec: prismatic cohomology}
Prismatic cohomology will be defined on the prismatic site associated to a formal scheme. For preliminaries on the latter, consult \cite[\href{https://stacks.math.columbia.edu/tag/0AHY}{Tag 0AHY}]{stacks-project}. The key result of the theory of formal schemes is that an $\FF_{p}$-scheme $X$, the formal completion functor produces a $p$-adic formal scheme isomorphic to $X$ allowing for cohomology computations to be done in the category of $p$-adic formal schemes.

Covers in the prismatic site will be given by $(p,I)$-completely faithfully flat prisms over the base as we now define. 
\begin{definition}[$I$-Completely Faithfully Flat Morphism]\label{def: completely faithfully flat morphism}
    Let $A$ be a commutative ring and $I\subseteq A$ an ideal. An $A$-module $M$ is $I$-completely faithfully flat if:
    \begin{enumerate}[label=(\roman*)]
        \item For all $I$-torsion modules $N$, $M\otimes^{L}_{A}N$ is a complex concentrated in degree 0 in $D(A)$. 
        \item $M\otimes^{L}_{A}(A/I)$ is a complex concentrated in degree 0 in $D(A)$ and faithfully flat as an $A/I$-module. 
    \end{enumerate}
\end{definition}
Given a $X$ smooth $p$-adic formal scheme over $\spf(A/I)$ with $(A,I)$ a bounded prism, we can construct the prismatic site $(X/A)_{\prism}$ as follows. 
\begin{definition}[Prismatic Site]\label{def: prismatic site}
    Let $X$ be a smooth $p$-adic formal scheme over $\spf(A/I)$ for some bounded prism $(A,I)$. The prismatic site $(X/A)_{\prism}$ is given by:
    \begin{itemize}
        \item Objects $\spf(B)\leftarrow \spf(B/IB)\to X$ where $(A,I)\to(B, IB)$ is a map of bounded prisms and $\spf(B/IB)\to X$ a map of $p$-adic formal schemes over $\spf(A/I)$. 
        \item Morphisms $(\spf(C)\rightarrow\spf(C/IC)\to X)\to(\spf(B)\leftarrow\spf(B/IB)\to X)$ where $(B,IB)\to (C,IC)$ is a map of bounded prisms under $(A, I)$ and  $\spf(C/IC)\to\spf(B/IB)$ is a map of formal schemes over $X$. 
        \item Covers $(\spf(C)\rightarrow\spf(C/IC)\to X)\to(\spf(B)\leftarrow\spf(B/IB)\to X)$ where $(B,IB)\to (C,IC)$ is a morphism of prisms such that $C$ is a $(p,IB)$-completely faithfully flat $B$-module. 
    \end{itemize}
\end{definition}
\begin{remark}
    If $X=\spf(R)$ then we denote the prismatic site $(R/A)_{\prism}$. 
\end{remark}
We can define presheaves $\Ocal_{\prism}$ and $\overline{\Ocal}_{\prism}$ by  
\begin{equation}\label{eqn: Oprism and Oprismbar}
    \begin{aligned}
        \Ocal_{\prism}\left(\spf(B)\leftarrow \spf(B/IB)\to X\right) &= B\\
        \overline{\Ocal}_{\prism}\left(\spf(B)\leftarrow \spf(B/IB)\to X\right) &= B/IB\\
    \end{aligned}
\end{equation}
which are in fact sheaves. 
\begin{proposition}[{\cite[Corr. 3.12]{BSPrismatic}}]\label{prop: Oprism and Oprismbar are sheaves}
    Let $X$ be a $p$-adic formal scheme over $\spf(A/I)$ for some bounded prism $(A,I)$. The presheaves $\Ocal_{\prism}$ and $\overline{\Ocal}_{\prism}$ on $(X/A)_{\prism}$ in (\ref{eqn: Oprism and Oprismbar}) are sheaves. 
\end{proposition}
These sheaves are known as the structure sheaf and reduced structure sheaf on the prismatic site. 
\begin{definition}[Structure Sheaves on the Prismatic Site]\label{def: structure sheaves on the prismatic stite}
    Let $X$ be a $p$-adic formal scheme over $\spf(A/I)$ for some bounded prism $(A,I)$. The sheaf $\Ocal_{\prism}$ (resp. $\overline{\Ocal}_{\prism}$) is the structure sheaf (resp. reduced structure sheaf) on the prismatic site. 
\end{definition}
Prismatic cohomology of a $p$-adic formal scheme is taken to be the derived functor cohomology on the prismatic site. 
\begin{definition}[Prismatic Cohomology]\label{def: prismatic cohomology}
    Let $X$ be a smooth formal scheme over $\spf(A/I)$ for some bounded prism $(A, I)$. The prismatic cohomology of $X/A$ is 
    $$R\Gamma_{\prism}(X/A)=R\Gamma((X/A)_{\prism},\Ocal_{\prism})\in D(A).$$
\end{definition}
\begin{remark}
    We define the analogous ringed site with reduced structure sheaf $\overline{\Ocal}_{\prism}$ the reduced prismatic site, which we denote $(X/A)_{\overline{\prism}}$.
\end{remark}
\begin{remark}
    This complex of $A$-modules is equipped with the $\widetilde{\varphi}_{A}$-linear endomorphism arising from the Frobenius action on $\Ocal_{\prism}$. 
\end{remark}