\documentclass{amsart}
\usepackage[margin=1.5in]{geometry}
\usepackage{amsmath}
\usepackage{tcolorbox}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{lastpage}
\usepackage{fancyhdr}
\usepackage{accents}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{color}
\usepackage{euscript}
\usepackage[bbgreekl]{mathbbol}
\DeclareSymbolFontAlphabet{\mathbb}{AMSb} % to ensure \mathbb does not change
\DeclareSymbolFontAlphabet{\mathbbl}{bbold}
\input{shortcuts.tex}
\setlength{\headheight}{40pt}


\newenvironment{solution}
  {\renewcommand\qedsymbol{$\blacksquare$}
  \begin{proof}[Solution]}
  {\end{proof}}
\renewcommand\qedsymbol{$\blacksquare$}

\usepackage{amsmath, amssymb, tikz, amsthm, csquotes, multicol, footnote, tablefootnote, biblatex, wrapfig, float, quiver, mathrsfs, cleveref, enumitem, upgreek, todonotes}
\addbibresource{refs.bib}
\theoremstyle{definition}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{exercise}[theorem]{Exercise}
\newtheorem{question}[theorem]{Question}
\newtheorem{fact}[theorem]{Fact}
\newtheorem{aside}[theorem]{Aside}
\newtheorem{example}[theorem]{Example}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{pseudodef}[theorem]{Pseudodefinition}
\newtheorem{pseudoprop}[theorem]{Pseudoproposition}
\numberwithin{equation}{section}

\setuptodonotes{color=blue!20, size=tiny}


\begin{document}
\large
\title[D\'{e}calage I]{The Inaugural Kleine AT \\ Talk 3: D\'{e}calage I\\ 10th October 2025}
\author{Wern Juin Gabriel Ong}
\address{Universit\"{a}t Bonn, Bonn, D-53113}
\email{wgabrielong@uni-bonn.de}
\urladdr{https://wgabrielong.github.io/}
\maketitle
\section*{Preliminaries}
These are the author's notes for the third talk at the inaugural Kleine AT conference on the construction of spectral sequences from filtered objects via d\'{e}calage.

\subsection*{Acknowledgements} The author also thanks his fellow organizer David Bowman for making this conference possible, and Mathieu Wydra for feedback on a preliminary version of this talk.
\large
\section{Recollections}\label{sec: recollections}
Recall the following from the first talk:
\begin{itemize}
  \item Let $\Cscr$ be a stable $\infty$-category. This is the appropriate analogue of an Abelian category in the higher setting (eg. fibers and cofibers exist).
  \item We will assume that $\Cscr$ has sequential limits and colimits.
  \item For a decreasing filtration $F^{\star}\in\Fil(\Cscr)\simeq\Fun(\ZZ^{\Opp},\Cscr)$, we denote $|F^{\star}|\simeq\colim_{s}F^{-s}$ its underlying object.
  \item The category $\Fil(\Cscr)$ of decreasing filtrations in $\Cscr$ is stable and can be endowed with the Beilinson $t$-structure.
\end{itemize}
\section{Spectral Sequences in General}\label{sec: sseq in general}
We first consider the simple case of a spectral sequence defined on an Abelian 1-category.
\begin{definition}[Spectral Sequence -- Abelian Category]\label{def: spectral sequence}
    Let $\Asf$ be an Abelian category. A spectral sequence in $\Asf$ starting at the $a$th page consists of:
    \begin{enumerate}[label=(\roman*)]
        \item Objects $E^{r}_{s,t}$ of $\Asf$ for $r\geq a$ and $s,t\in\ZZ$.
        \item Differentials $d^{r}_{s,t}:E^{r}_{s,t}\to E^{r}_{s-r,t+r-1}$ for $r\geq a$ and $s,t\in\ZZ$.
        \item Isomorphisms between the cohomology at $E^{r}_{s,t}$ with respect to $d^{r}$ and $E^{r+1}_{s,t}$ for all $r\geq a$ and $s,t\in\ZZ$.
    \end{enumerate}
\end{definition}
We will by convention assume that spectral sequences start at page 1. So the $E^{1}$-page of a spectral sequence is given by
\begin{figure}[H]\label{fig: E1 page of sseq}
  $$% https://q.uiver.app/#q=WzAsMTUsWzIsMSwiRV57MX1fezAsMH0iXSxbMywxLCJFXnsxfV97MSwwfSJdLFsxLDEsIkVeezF9X3stMSwwfSJdLFswLDEsIkVeezF9X3stMiwwfSJdLFswLDAsIkVeezF9X3stMiwxfSJdLFsxLDAsIkVeezF9X3stMSwxfSJdLFsyLDAsIkVeezF9X3swLDF9Il0sWzMsMCwiRV57MX1fezEsMX0iXSxbNCwwLCJFXnsxfV97MiwxfSJdLFs0LDEsIkVeezF9X3syLDB9Il0sWzQsMiwiRV57MX1fezIsLTF9Il0sWzMsMiwiRV57MX1fezEsLTF9Il0sWzIsMiwiRV57MX1fezAsLTF9Il0sWzEsMiwiRV57MX1fey0xLC0xfSJdLFswLDIsIkVeezF9X3stMiwtMX0iXSxbOCw3LCJkXnsxfV97MiwxfSIsMl0sWzcsNiwiZF57MX1fezEsMX0iLDJdLFs2LDUsImReezF9X3swLDF9IiwyXSxbNSw0LCJkXnsxfV97LTEsMX0iLDJdLFsyLDMsImReezF9X3stMSwwfSIsMl0sWzAsMiwiZF57MX1fezAsMH0iLDJdLFsxLDAsImReezF9X3sxLDB9IiwyXSxbOSwxLCJkXnsxfV97MiwwfSIsMl0sWzEwLDExLCJkXnsxfV97MiwtMX0iLDJdLFsxMSwxMiwiZF57MX1fezEsLTF9IiwyXSxbMTIsMTMsImReezF9X3swLC0xfSIsMl0sWzEzLDE0LCJkXnsxfV97LTEsLTF9IiwyXV0=
  \begin{tikzcd}
    {E^{1}_{-2,1}} & {E^{1}_{-1,1}} & {E^{1}_{0,1}} & {E^{1}_{1,1}} & {E^{1}_{2,1}} \\
    {E^{1}_{-2,0}} & {E^{1}_{-1,0}} & {E^{1}_{0,0}} & {E^{1}_{1,0}} & {E^{1}_{2,0}} \\
    {E^{1}_{-2,-1}} & {E^{1}_{-1,-1}} & {E^{1}_{0,-1}} & {E^{1}_{1,-1}} & {E^{1}_{2,-1}}
    \arrow["{d^{1}_{-1,1}}"', from=1-2, to=1-1]
    \arrow["{d^{1}_{0,1}}"', from=1-3, to=1-2]
    \arrow["{d^{1}_{1,1}}"', from=1-4, to=1-3]
    \arrow["{d^{1}_{2,1}}"', from=1-5, to=1-4]
    \arrow["{d^{1}_{-1,0}}"', from=2-2, to=2-1]
    \arrow["{d^{1}_{0,0}}"', from=2-3, to=2-2]
    \arrow["{d^{1}_{1,0}}"', from=2-4, to=2-3]
    \arrow["{d^{1}_{2,0}}"', from=2-5, to=2-4]
    \arrow["{d^{1}_{-1,-1}}"', from=3-2, to=3-1]
    \arrow["{d^{1}_{0,-1}}"', from=3-3, to=3-2]
    \arrow["{d^{1}_{1,-1}}"', from=3-4, to=3-3]
    \arrow["{d^{1}_{2,-1}}"', from=3-5, to=3-4]
  \end{tikzcd}$$
  \caption{$E^{1}$-page of a spectral sequence.}
\end{figure}
where all the rows are chain complexes in $\Asf$. We take the homology at each entry to get to the $E^{2}$-page which is given by
\begin{figure}[H]\label{fig: E2 page of sseq}
  $$% https://q.uiver.app/#q=WzAsMTUsWzIsMSwiRV57Mn1fezAsMH0iXSxbMywxLCJFXnsyfV97MSwwfSJdLFsxLDEsIkVeezJ9X3stMSwwfSJdLFswLDEsIkVeezJ9X3stMiwwfSJdLFswLDAsIkVeezJ9X3stMiwxfSJdLFsxLDAsIkVeezJ9X3stMSwxfSJdLFsyLDAsIkVeezJ9X3swLDF9Il0sWzMsMCwiRV57Mn1fezEsMX0iXSxbNCwwLCJFXnsyfV97MiwxfSJdLFs0LDEsIkVeezJ9X3syLDB9Il0sWzQsMiwiRV57Mn1fezIsLTF9Il0sWzMsMiwiRV57Mn1fezEsLTF9Il0sWzIsMiwiRV57Mn1fezAsLTF9Il0sWzEsMiwiRV57MX1fey0xLC0xfSJdLFswLDIsIkVeezJ9X3stMiwtMX0iXSxbMTAsMCwiZF57Mn1fezIsLTF9IiwxXSxbMTEsMiwiZF57Mn1fezEsLTF9IiwxXSxbMTIsMywiZF57Mn1fezAsLTF9IiwxXSxbMCw0LCJkXnsyfV97MCwwfSIsMV0sWzEsNSwiZF57Mn1fezEsMH0iLDFdLFs5LDYsImReezJ9X3syLDB9IiwxXV0=
  \begin{tikzcd}
    {E^{2}_{-2,1}} & {E^{2}_{-1,1}} & {E^{2}_{0,1}} & {E^{2}_{1,1}} & {E^{2}_{2,1}} \\
    {E^{2}_{-2,0}} & {E^{2}_{-1,0}} & {E^{2}_{0,0}} & {E^{2}_{1,0}} & {E^{2}_{2,0}} \\
    {E^{2}_{-2,-1}} & {E^{1}_{-1,-1}} & {E^{2}_{0,-1}} & {E^{2}_{1,-1}} & {E^{2}_{2,-1}}
    \arrow["{d^{2}_{0,0}}"{description}, from=2-3, to=1-1]
    \arrow["{d^{2}_{1,0}}"{description}, from=2-4, to=1-2]
    \arrow["{d^{2}_{2,0}}"{description}, from=2-5, to=1-3]
    \arrow["{d^{2}_{0,-1}}"{description}, from=3-3, to=2-1]
    \arrow["{d^{2}_{1,-1}}"{description}, from=3-4, to=2-2]
    \arrow["{d^{2}_{2,-1}}"{description}, from=3-5, to=2-3]
  \end{tikzcd}$$
  \caption{$E^{2}$-page of a spectral sequence.}
\end{figure}
\noindent where we have implicitly supplied isomorphisms,
$$E^{2}_{s,t}\cong\frac{\ker\left(d^{1}_{s,t}:E^{1}_{s,t}\to E^{1}_{s-1,t}\right)}{\img\left(d^{1}_{s+1,t}:E^{1}_{s+1,t}\to E^{1}_{s,t}\right)}.$$
Spectral sequences are an important computational tool: for a difficult-to-understand object $X$, we can populate $E^{a}_{s,t}$ with some computable invariants of $X$ and hope to understand $X$ better from the terms of $E^{r}_{s,t}$. This is illustrated by the following example.
\begin{example}[Atiyah-Hirzebruch Spectral Sequence]\label{ex: AHSS}
  Let $\Hcal$ be a nice cohomology theory and $X$ a CW complex. The Atiyah-Hirzebruch spectral sequence is the spectral sequence starting on the $E^{2}$-page with
  $$E^{2}_{s,t}=H^{s}(X,\Hcal^{t}(\{*\})).$$
  It is a fact that the Atiyah-Hirzebruch spectral sequence stabilizes for sufficiently large $r$ (ie. the $E^{r}$ and $E^{r+1}$-pages agree for $r>>>0$), and it is in this sense that a spectral sequence can be said to converge. The so-called $E^{\infty}$-page produces the associated grades for a filtration on $\Hcal^{s+t}(X)$ along the diagonals. The terms of the $E^{2}$-page are easily obtained by the universal coefficient theorem.
\end{example}
More generally, for a stable $\infty$-category, we know that its heart $\Cscr^{\heartsuit}$ is an Abelian 1-category, and \Cref{def: spectral sequence} generalizes immediately to $\Cscr$ in the obvious way.
\begin{definition}[Spectral Sequence -- Stable Category]\label{def: spectral sequence stable}
    Let $\Cscr$ be a stable $\infty$-category. A spectral sequence in $\Cscr$ starting at the $a$th page consists of:
    \begin{enumerate}[label=(\roman*)]
        \item Objects $E^{r}_{s,t}$ of $\Cscr^{\heartsuit}$ for $r\geq a$ and $s,t\in\ZZ$.
        \item Differentials $d^{r}_{s,t}:E^{r}_{s,t}\to E^{r}_{s-r,t+r-1}$ for $r\geq a$ and $s,t\in\ZZ$.
        \item Isomorphisms between the cohomology at $E^{r}_{s,t}$ with respect to $d^{r}$ and $E^{r+1}_{s,t}$ for all $r\geq a$ and $s,t\in\ZZ$.
    \end{enumerate}
\end{definition}
\section{Spectral Sequence of a Filtered Object}\label{sec: sseq of filtered object}
One way that spectral sequences arises is from filtered objects, the desideratum here being that we want to make conclusions about the homotopy objects $\pi_{*}|F|$ of the underlying object $|F|$ for $F\in\Fil(\Cscr)$ from the associated graded objects which are more easily acccessible.

Observe here that a spectral sequence (which we have assumed to start at the $E^{1}$-page) has chain complexes along the horizontals with differentials pointing left.

Recalling the result from the first talk that the composition $\gr_{F}^{s}\to\gr_{F}^{s+1}[1]\to\gr_{F}^{s+2}[2]$ is canonically nullhomotopic \cite[Lem. 3.18]{Antieau}, allows us to think of the sequential diagram
\begin{equation}\label{eqn: cochain complex on associated graded}
  \dots\to\gr_{F}^{-s-1}[-s-1]\to\gr_{F}^{-s}[-s]\to\gr_{F}^{-s+1}[-s+1]\to\dots
\end{equation}
as a cochain complex in $\Cscr$. After passing to homotopy objects -- and appealing to commutativity of the formation of homotopy objects with the shift -- we have that
\begin{equation}\label{eqn: cochain complex on associated graded heart}
  \dots\to\pi_{s+t+1}\gr_{F}^{-s-1}\to\pi_{s+t}\gr_{F}^{-s}\to\pi_{s+t-1}\gr_{F}^{-s+1}\to\dots
\end{equation}
is an honest cochain complex in $\Cscr^{\heartsuit}$. The cochain complexes (\ref{eqn: cochain complex on associated graded heart}) form natural candidates for the horizonal entries of a spectral sequence that we could associate to $F$.

We will show this expectation indeed holds.
\begin{proposition}\label{prop: spectral sequence of filtered object}
  Let $\Cscr$ be a stable $\infty$-category and $F^{\star}$ decreasing filtration in $\Cscr$. Define for $r\geq 1$ the collection of objects
  $$E^{r}_{s,t}=\img\left(\pi_{s+t}\gr_{F}^{[-s,-s+r)}\to\pi_{s+t}\gr_{F}^{[-s-r+1,-s+1)}\right)\in\Cscr^{\heartsuit}.$$
  There exist differentials $d^{r}_{s,t}:E^{r}_{s,t}\to E^{r}_{s-r, t+r-1}$ and isomorphisms between the cohomology at $E^{r}_{s,t}$ and $E^{r+1}_{s,t}$ such that the objects, together with the differentials and isomorphisms, define a spectral sequence in the sense of \Cref{def: spectral sequence stable} starting on the $E^{1}$-page. In particular, the $E^{1}$-page of this spectral sequence is given by $E^{1}_{s,t}=\pi_{s+t}\gr^{-s}_{F}$, and the differentials agree with those in (\ref{eqn: cochain complex on associated graded heart}).
\end{proposition}
\begin{proof}
  For the first part, we only show that the maps $d^{r}_{s,t}:E^{r}_{s,t}\to E^{r}_{s-r, t+r-1}$ exist. That these maps are indeed differentials and the isomorphism from homology to the term on the successive page will follow from the comparison that arises from the d\'{e}calage construction.

  Using the commuting diagram
  $$% https://q.uiver.app/#q=WzAsNixbMCwwLCJGXnstcysycn0iXSxbMiwwLCJGXnstcytyfSJdLFs0LDAsIkZeey1zfSJdLFs0LDEsIkZeey1zLXIrMX0iXSxbMiwxLCJGXnstcysxfSJdLFswLDEsIkZeey1zK3IrMX0iXSxbMCw1XSxbMCwxXSxbMSwyXSxbMiwzXSxbNCwzXSxbNSw0XSxbMSw0XV0=
  \begin{tikzcd}
    {F^{-s+2r}} && {F^{-s+r}} && {F^{-s}} \\
    {F^{-s+r+1}} && {F^{-s+1}} && {F^{-s-r+1}}
    \arrow[from=1-1, to=1-3]
    \arrow[from=1-1, to=2-1]
    \arrow[from=1-3, to=1-5]
    \arrow[from=1-3, to=2-3]
    \arrow[from=1-5, to=2-5]
    \arrow[from=2-1, to=2-3]
    \arrow[from=2-3, to=2-5]
  \end{tikzcd}$$
  we obtain by an iterated taking of cofibers the diagram
  $$% https://q.uiver.app/#q=WzAsNixbMCwwLCJcXGdyX3tGfV57Wy1zK3IsLXMrMnIpfSJdLFsyLDAsIlxcZ3Jfe0Z9XntbLXMsLXMrMnIpfSJdLFs0LDAsIlxcZ3Jfe0Z9XntbLXMsLXMrcil9Il0sWzAsMSwiXFxncl97Rn1ee1stcysxLC1zK3IrMSl9Il0sWzIsMSwiXFxncl97Rn1ee1stcy1yKzEsLXMrcisxKX0iXSxbNCwxLCJcXGdyX3tGfV57Wy1zLXIrMSwtcysxKX0iXSxbMCwxXSxbMSwyXSxbMiw1XSxbMSw0XSxbMCwzXSxbMyw0XSxbNCw1XV0=
  \begin{tikzcd}
    {\gr_{F}^{[-s+r,-s+2r)}} && {\gr_{F}^{[-s,-s+2r)}} && {\gr_{F}^{[-s,-s+r)}} \\
    {\gr_{F}^{[-s+1,-s+r+1)}} && {\gr_{F}^{[-s-r+1,-s+r+1)}} && {\gr_{F}^{[-s-r+1,-s+1)}}
    \arrow[from=1-1, to=1-3]
    \arrow[from=1-1, to=2-1]
    \arrow[from=1-3, to=1-5]
    \arrow[from=1-3, to=2-3]
    \arrow[from=1-5, to=2-5]
    \arrow[from=2-1, to=2-3]
    \arrow[from=2-3, to=2-5]
  \end{tikzcd}$$
  with rows cofiber sequences. Passing to homotopy objects gives the commutative diagram of dotted arrows
  $${\tiny
 % https://q.uiver.app/#q=WzAsMTAsWzEsMiwiXFxwaV97cyt0fVxcZ3Jfe0Z9XntbLXMtcisxLC1zKzEpfSJdLFsxLDAsIlxccGlfe3MrdH1cXGdyX3tGfV57Wy1zLC1zK3IpfSJdLFs0LDAsIlxccGlfe3MrdC0xfVxcZ3Jfe0Z9XntbLXMrciwtcysycil9Il0sWzQsMiwiXFxwaV97cyt0LTF9XFxncl97Rn1ee1stcysxLC1zK3IrMSl9Il0sWzIsMSwiRV57cn1fe3MsdH0iXSxbMywxLCJFXntyfV97cy1yLHQrci0xfSJdLFswLDIsIlxcZG90cyJdLFswLDAsIlxcZG90cyJdLFs1LDIsIlxcZG90cyJdLFs1LDAsIlxcZG90cyJdLFsxLDIsIiIsMSx7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6ImRvdHRlZCJ9fX1dLFsxLDAsIiIsMSx7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6ImRvdHRlZCJ9fX1dLFsyLDMsIiIsMSx7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6ImRvdHRlZCJ9fX1dLFswLDMsIiIsMSx7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6ImRvdHRlZCJ9fX1dLFsyLDVdLFs1LDNdLFsxLDRdLFs0LDBdLFs0LDUsIlxcZXhpc3RzISIsMCx7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6ImRhc2hlZCJ9fX1dLFs2LDAsIiIsMCx7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6ImRvdHRlZCJ9fX1dLFs3LDEsIiIsMCx7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6ImRvdHRlZCJ9fX1dLFszLDgsIiIsMCx7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6ImRvdHRlZCJ9fX1dLFsyLDksIiIsMCx7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6ImRvdHRlZCJ9fX1dXQ==
  \begin{tikzcd}
    \dots & {\pi_{s+t}\gr_{F}^{[-s,-s+r)}} &&& {\pi_{s+t-1}\gr_{F}^{[-s+r,-s+2r)}} & \dots \\
    && {E^{r}_{s,t}} & {E^{r}_{s-r,t+r-1}} \\
    \dots & {\pi_{s+t}\gr_{F}^{[-s-r+1,-s+1)}} &&& {\pi_{s+t-1}\gr_{F}^{[-s+1,-s+r+1)}} & \dots
    \arrow[dotted, from=1-1, to=1-2]
    \arrow[dotted, from=1-2, to=1-5]
    \arrow[from=1-2, to=2-3]
    \arrow[dotted, from=1-2, to=3-2]
    \arrow[dotted, from=1-5, to=1-6]
    \arrow[from=1-5, to=2-4]
    \arrow[dotted, from=1-5, to=3-5]
    \arrow["{\exists!}", dashed, from=2-3, to=2-4]
    \arrow[from=2-3, to=3-2]
    \arrow[from=2-4, to=3-5]
    \arrow[dotted, from=3-1, to=3-2]
    \arrow[dotted, from=3-2, to=3-5]
    \arrow[dotted, from=3-5, to=3-6]
  \end{tikzcd}
  \normalsize}$$
  and the rows are the long exact sequence for homotopy objects of a fiber sequence. The solid arrows exist by factorization of the vertical maps into an epimorphism followed by a monomorphism, and the dashed arrow exists uniquely by functoriality of such factorizations.

  For the second statement, setting $r=1$, we have that
  \begin{align*}
    E^{1}_{s,t} &\cong \img\left(\pi_{s+t}\gr_{F}^{[-s,-s+1)}\to\pi_{s+t}\gr_{F}^{[-s,s+1)}\right) \\
    &\cong \img\left(\pi_{s+t}\gr_{F}^{-s}\xrightarrow{\id_{\pi_{s+t}\gr_{F}^{-s}}}\pi_{s+t}\gr_{F}^{-s}\right) \\
    &\cong \pi_{s+t}\gr_{F}^{-s}
  \end{align*}
  where in the middle we have used that $\gr_{F}^{[-s,-s+1)}\simeq\cofib(F^{-s+1}\to F^{-s})\simeq\gr^{i}_{F}$ and that the map is induced along the identity. Compatibility with the differentials is clear as unwinding the preceding constructions for $r=1$ we recover precisely the map $\delta:\gr_{F}^{-s}\to\gr_{F}^{-s+1}[1]$ of the previous talk (cf. \cite[Lem. 3.18]{Antieau}).
\end{proof}
\begin{definition}[Lurie -- Spectral Sequence of Filtered Object]\label{def: sseq of filtered object}
  Let $\Cscr$ be a stable $\infty$-category and $F^{\star}$ decreasing filtration in $\Cscr$. The (Lurie) spectral sequence $E^{r}_{s,t}(F)$ of $F^{\star}$ is the spectral sequence of \Cref{prop: spectral sequence of filtered object}.
\end{definition}
\begin{figure}[H]\label{fig: E1 page of sseq to filtered object}
  $$% https://q.uiver.app/#q=WzAsMTUsWzIsMSwiXFxwaV97MH1cXGdyX3tGfV57MH0iXSxbMywxLCJcXHBpX3sxfVxcZ3Jfe0Z9XnstMX0iXSxbNCwxLCJcXHBpX3syfVxcZ3Jfe0Z9XnstMn0iXSxbMCwxLCJcXHBpX3stMn1cXGdyX3tGfV57Mn0iXSxbMSwxLCJcXHBpX3stMX1cXGdyX3tGfV57MX0iXSxbMCwwLCJcXHBpX3stMX1cXGdyX3tGfV57Mn0iXSxbMSwwLCJcXHBpX3swfVxcZ3Jfe0Z9XnsxfSJdLFsyLDAsIlxccGlfezF9XFxncl97Rn1eezB9Il0sWzMsMCwiXFxwaV97Mn1cXGdyX3tGfV57LTF9Il0sWzQsMCwiXFxwaV97M31cXGdyX3tGfV57LTJ9Il0sWzAsMiwiXFxwaV97LTN9XFxncl97Rn1eezJ9Il0sWzEsMiwiXFxwaV97LTJ9XFxncl97Rn1eezF9Il0sWzIsMiwiXFxwaV97LTF9XFxncl97Rn1eezB9Il0sWzMsMiwiXFxwaV97MH1cXGdyX3tGfV57LTF9Il0sWzQsMiwiXFxwaV97MX1cXGdyX3tGfV57LTJ9Il0sWzE0LDEzXSxbMTMsMTJdLFsxMiwxMV0sWzExLDEwXSxbMiwxXSxbMSwwXSxbMCw0XSxbNCwzXSxbOSw4XSxbOCw3XSxbNyw2XSxbNiw1XV0=
  \begin{tikzcd}
    {\pi_{-1}\gr_{F}^{2}} & {\pi_{0}\gr_{F}^{1}} & {\pi_{1}\gr_{F}^{0}} & {\pi_{2}\gr_{F}^{-1}} & {\pi_{3}\gr_{F}^{-2}} \\
    {\pi_{-2}\gr_{F}^{2}} & {\pi_{-1}\gr_{F}^{1}} & {\pi_{0}\gr_{F}^{0}} & {\pi_{1}\gr_{F}^{-1}} & {\pi_{2}\gr_{F}^{-2}} \\
    {\pi_{-3}\gr_{F}^{2}} & {\pi_{-2}\gr_{F}^{1}} & {\pi_{-1}\gr_{F}^{0}} & {\pi_{0}\gr_{F}^{-1}} & {\pi_{1}\gr_{F}^{-2}}
    \arrow[from=1-2, to=1-1]
    \arrow[from=1-3, to=1-2]
    \arrow[from=1-4, to=1-3]
    \arrow[from=1-5, to=1-4]
    \arrow[from=2-2, to=2-1]
    \arrow[from=2-3, to=2-2]
    \arrow[from=2-4, to=2-3]
    \arrow[from=2-5, to=2-4]
    \arrow[from=3-2, to=3-1]
    \arrow[from=3-3, to=3-2]
    \arrow[from=3-4, to=3-3]
    \arrow[from=3-5, to=3-4]
  \end{tikzcd}$$
  \caption{$E^{1}$-page of the spectral sequence of a filtered object.}
\end{figure}
It is easy to see that the $t$-th row of $E^{1}_{s,t}(F)$ is precisely the cochain complex (\ref{eqn: cochain complex on associated graded heart}).
\section{D\'{e}calage}\label{sec: decalage}
The d\'{e}calage construction produces a new filtered object $\Dec(F)\in\Fil(\Cscr)$ in a natural way from the Beilinson $t$-structure on $\Cscr$.

Recall that formation of Whitehead towers $\tau_{\geq n}^{B}:\Fil(\Cscr)\to\Fil(\Cscr)_{\geq n}$ is right adjoint to the full inclusion of $\Fil(\Cscr)_{\geq n}\subseteq\Fil(\Cscr)$. This furnishes natural maps $\tau_{\geq n}^{B}F\to F^{\star}$ and more generally $\tau_{\geq n+1}^{B}F\to\tau_{\geq n}^{B}F$ in $\Fil(\Cscr)$ which compose to the sequential diagram
\begin{equation}\label{eqn: dec before realization}
  \dots\to\tau_{\geq n+1}^{B}F\to\tau_{\geq n}^{B}F\to\tau_{\geq n-1}^{B}F\to\dots
\end{equation}
and the d\'{e}calage of $F$ is defined to be the filtered object obtained by passing to the realization of (\ref{eqn: dec before realization}).
\begin{definition}[D\'{e}calage]\label{def: decalage}
    Let $\Cscr$ be a stable $\infty$-category and $F^{\star}$ a decreasing filtration in $\Cscr$. The d\'{e}calage $\Dec(F)^{\star}$ of $F$ is the decreasing filtration in $\Cscr$ given by $\Dec(F)^{i}=|\tau_{\geq i}^{B}F|$.
\end{definition}
Here, we have treated each Whitehead tower $\tau_{\geq n}^{B}F$ as a filtered object, and considered its underlying object.

Before going further, let us verify record elementary properties of the d\'{e}calage.
\begin{lemma}\label{lem: properties of dec}
  Let $\Cscr$ be a stable $\infty$-category and $F^{\star}$ a decreasing filtration in $\Cscr$.
  \begin{enumerate}[label=(\roman*)]
    \item There is an equivalence on underlying objects $|F^{\star}|\simeq|\Dec(F)^{\star}|$.
    \item The associated graded $\gr^{n}_{\Dec(F)}$ is given by $|\pi_{n}^{B}F|[n]\simeq|(\pi_{n}^{B}F)[n]|$, where $\pi_{n}^{B}F$ is the cochain complex
    $$\dots\to\pi_{n+1}\gr_{F}^{-1}\to\pi_{n}\gr_{F}^{0}\to\pi_{n-1}\gr_{F}^{1}\to\dots$$
    in $\Cscr^{\heartsuit}$ where $\pi_{n}\gr_{F}^{0}$ is in cohomological degree 0.
  \end{enumerate}
\end{lemma}
\begin{proof}[Proof of (i)]
  This is clear from $\tau_{\geq-\infty}^{B}F\simeq F^{\star}$.
\end{proof}
\begin{proof}[Proof of (ii)]
  We compute
  \begin{align*}
    \gr^{n}_{\Dec(F)} &\simeq \cofib\left(\left|\tau_{\geq n+1}^{B}F\right|\to\left|\tau_{\geq n}^{B}F\right|\right) \\
    &\simeq \left|\cofib\left(\tau_{\geq n+1}^{B}F\to\tau_{\geq n}^{B}F\right)\right| \\
    &\simeq \left|(\pi_{n}^{B}F)[n]\right|
  \end{align*}
  and using that the shift is computed pointwise gives the first statement.

  For the second statement, we use that $\pi_{n}^{B}F\in\Fil(\Cscr)^{\heartsuit}$ so we can use the equivalence $\Ch^{\bullet}(\Cscr^{\heartsuit})\simeq\Fil(\Cscr)^{\heartsuit}$ to identify $\pi_{n}^{B}F$ with the cochain complex $C^{\bullet}$ with $C^{i}=\gr^{i}_{\pi_{n}^{B}F}[i]$. By $t$-exactness of the associated graded, we have that $\gr^{i}_{\pi_{n}^{B}F}[i]\simeq \pi_{n-i}\gr_{F}^{i}$. See \cite[Prop. 5.8 \& Corr. 5.9]{Ariotta} for more details.
\end{proof}
Denote $\Dec^{(r)}(F)$ to be the $r$-fold iterate of the d\'{e}calage functor, taking $\Dec^{(0)}(F)$ to be the identity. The preceding lemma shows that the underlying objects $|F^{\star}|$ and $|\Dec^{(r)}(F)^{\star}|$ agree.
\begin{lemma}\label{lem: decalage spectral sequence}
  Let $\Cscr$ be a stable $\infty$-category and $F^{\star}$ a decreasing filtration in $\Cscr$. Define for $r\geq 0$ the collection of objects
  $$E^{r+1}_{s,t}=\pi_{s+t}\gr_{\Dec^{(r)}(F)}^{rs+rt-s}.$$
  There exist differentials $d^{r}_{s,t}:E^{r}_{s,t}\to E^{r}_{s-r, t+r-1}$ and isomorphisms between the cohomology at $E^{r}_{s,t}$ and $E^{r+1}_{s,t}$ such that the objects, together with the differentials and isomorphisms, define a spectral sequence in the sense of \Cref{def: spectral sequence stable} starting on the $E^{1}$-page. In particular, the $E^{1}$-page of this spectral sequence is given by $E^{1}_{s,t}=\pi_{s+t}\gr^{-s}_{F}$, and the differentials agree with those in (\ref{eqn: cochain complex on associated graded heart}).
\end{lemma}
\begin{proof}
  The second statement just follows formally from the fact that the $0$-fold composite of the d\'{e}calage functor is the identity on $F$.

  By induction, it suffices to consider the case of $r=1$. Here we have maps $d^{2}_{s,t}:\pi_{s+t}\gr_{\Dec(F)}^{t}\to\pi_{s+t-1}\gr_{\Dec(F)}^{t-1}$ but we already know these are differentials. The homology at $E^{1}_{s,t}$ is given by
  \begin{align*}
    \frac{\ker\left(\pi_{s+t}\gr^{-s}_{F}\to\pi_{s+t-1}\gr_{F}^{-s+1}\right)}{\img\left(\pi_{s+t+1}\gr_{F}^{-s-1}\to\pi_{s+t}\gr_{F}^{-s}\right)} &= H^{-s}(\pi_{t}^{B}F)\\
    &\cong \pi_{s}|\pi_{t}^{B}F| \\
    &= \pi_{s+t}|\pi_{t}^{B}F|[t]\\
    &= \pi_{s+t}\gr_{\Dec(F)}^{t} = E^{2}_{s,t}
  \end{align*}
  where the first equality is by the definition of $\pi_{t}^{B}F$ as the cochain complex
  $$% https://q.uiver.app/#q=WzAsOCxbMywwLCItcyJdLFsxLDAsIi1zKzEiXSxbNSwwLCItcy0xIl0sWzAsMSwiXFxkb3RzIl0sWzEsMSwiXFxwaV97cyt0KzF9XFxncl97Rn1eey1zLTF9Il0sWzMsMSwiXFxwaV97cyt0fVxcZ3Jfe0Z9Xnstc30iXSxbNSwxLCJcXHBpX3tzK3QtMX1cXGdyX3tGfV57LXMrMX0iXSxbNiwxLCJcXGRvdHMiXSxbMyw0XSxbNCw1XSxbNSw2XSxbNiw3XV0=
  \begin{tikzcd}
    & {-s+1} && {-s} && {-s-1} \\
    \dots & {\pi_{s+t+1}\gr_{F}^{-s-1}} && {\pi_{s+t}\gr_{F}^{-s}} && {\pi_{s+t-1}\gr_{F}^{-s+1}} & \dots
    \arrow[from=2-1, to=2-2]
    \arrow[from=2-2, to=2-4]
    \arrow[from=2-4, to=2-6]
    \arrow[from=2-6, to=2-7]
  \end{tikzcd}$$
  after reindexing, and the second canonical isomorphism provided by $\pi_{s}|C^{\bullet}|\cong H^{-s}(C^{\bullet})$ for cochain complexes $C^{\bullet}$.
\end{proof}
We can then define this as a spectral sequence.
\begin{definition}[D\'{e}calage -- Spectral Sequence of Filtered Object]\label{def: sseq of filtered object decalage}
  Let $\Cscr$ be a stable $\infty$-category and $F^{\star}$ decreasing filtration in $\Cscr$. The d\'{e}calage spectral sequence $E^{r}_{s,t}(F)$ of $F^{\star}$ is the spectral sequence of \Cref{lem: decalage spectral sequence}.
\end{definition}
\section{Towards the Comparison Result}\label{sec: comparison}
We have thus far produced two constructions for a spectral sequence associated to $F^{\star}\in\Fil(\Cscr)$
\begin{itemize}
  \item (Lurie) $E^{r}_{s,t}=\img\left(\pi_{s+t}\gr_{F}^{[-s,-s+r)}\to\pi_{s+t}\gr_{F}^{[-s-r+1,-s+1)}\right)$, and
  \item (D\'{e}calage) $E^{r}_{s,t}=\pi_{s+t}\gr_{\Dec^{(r-1)}(F)}^{rs+rt-2s-t}$.
\end{itemize}
We know that the $E^{1}$-pages of these two constructions agree, and by definition of a spectral sequence that the higher terms of the pages of the respective spectral sequences are abstractly isomorphic, albeit not necessarily in a way compatible with the differentials.

However, it turns out that these definitions are equivalent, and that taking d\'{e}calage computes the subsequent page of the spectral sequence (modulo some index shifting). We conclude by stating the result that will be shown in the following talk.
\begin{theorem}\label{thm: comparison}
  Let $\Cscr$ be a stable $\infty$-category and $F^{\star}$ decreasing filtration in $\Cscr$. For each $r\geq 1$ there are natural isomorphisms $E^{r}_{-t,s+2t}(\Dec(F))\cong E^{r+1}_{s,t}(F)$ compatible with the differentials.
\end{theorem}
\printbibliography
\end{document}
